---
title: "Project 3"
output: html_notebook
---

# PART 1

**
The data shown in the next page were collected by Spicer et al. (1987) in an
investigation of sudden infant death syndrome (SIDS). The two groups here consist of 16
SIDS victims and 49 controls. The Factor68 variable arises from spectral analysis of 24
hour recordings of electrocardiograms and respiratory movements made on each child.
All the infants have a gestational age of 37 weeks or more and were regarded as full
term. Enter the data in a Note Pad and save it as SIDS.txt
**

```{r}
library(MASS)
file <- "C:\\Users\\Taterthot\\Desktop\\da 410\\a4\\babydeath.txt"
sids <- read.table(file, header = TRUE, sep = "", dec = ".")
sids

```

# PART 2

**
Use R to construct the discriminant function based on all four variables and use it to classify a baby with HR = 100, BW = 3000, Factor68 = 0.3, Gesage = 40. 

Make sure you include clear command lines and relevant output/results with hypotheses, test result(s) and conclusion(s)/interpretation(s).
**

```{r}
sids.lda <- lda(Group~ ., data = sids,  prior=c(0.5,0.5))
sids.lda
```

```{r}
classobs <- rbind(c(100, 3000, 0.30, 40))
dimnames(classobs) <- list(NULL,c('HR','BW', 'Factor68', 'Gesage'))
classobs <- data.frame(classobs)
classobs
```

## PREDICT GROUP

```{r}
predict(sids.lda,newdata=classobs)$class
```

With the prediction we can say that a baby with HR = 100, BW=3000, Factor68=0.3, and Gesage=40 would classify with the sids of group 1.

## POSTERIOR PROBABILITIES

```{r}
predict(sids.lda,newdata=classobs)$posterior
```

With the likley-hood of prior probability and new evidence we can suggest that with the current baby prediction, 65.71% of sids babies will fall into group 1, while the other 34.29% would fall into group 2. We have further analysis of misclassification and cross-validation rates below.

## MISCLASSIFICATION RATE

```{r}
group<-predict(sids.lda, sids, method='plug-in')$class
table(group, sids$Group)
```

Plug-in missclassicifaction: (3+8) / (41+3+8+13) = 11/65 = 0.169.

## CROSS VALIDATION

```{r}
correct<-rep(0,times=nrow(sids) )
for (j in 1:nrow(sids) ) {
mydis<-lda(grouping=sids$Group[-j], x=sids[-j,2:5], prior=c(0.5,0.5))
mypred<-predict(mydis,newdata=sids[j,2:5])$class
correct[j] <- (mypred==sids$Group[j])
}
cv.misclass <- 1-mean(correct)
cv.misclass
```

Cross-validation misclassification rate for LDA is 0.20.