---
title: "Project 7"
output: html_notebook
---

## JONATHAN FENG

**
13.8) Use the ramus bone data of Table 3.7
**

```{r}
bone <- read.csv("C:\\Users\\Taterthot\\Desktop\\da 410\\a8\\bone.csv", fileEncoding = 'UTF-8-BOM')
bone
```

**
(a) Extract loadings by the principal component method and do a varimax rotation. Use two Factors.
**

# FINDING CORRELATION MATRIX C

```{r}
bone.cor <- cor(bone[,1:4])
bone.cor
```

# FINDING EIGENVALUE D AND EIGENVECTOR C

```{r}
eigen <- eigen(bone.cor)
eigen
```

```{r}
c <- eigen$vectors[,1:2]
c

d <- diag(eigen$values[1:2])
d
```

## LOADINGS LAMBDA 1 AND 2

```{r}
a <- c %*% sqrt(d)
a <- -(a)
round(a, 3)
```

## COMMUNALITIES

```{r}
com <- (a[,1]^2) + (a[,2]^2)
com <- round(com, 3)
com
```

## SPECIFIC VARIANCES

```{r}
var <- round((1 - (com)), 3)
var
```

## TOTAL VARIANCE ACCOUNTED FOR

```{r}
total.var <- eigen$values[1:2]
total.var
```

## PROPORTION OF TOTAL VARIANCE

```{r}
x <- rep(NA, 2)
for (i in 1:2) {
  x[i] <-
    eigen$values[i] / (eigen$values[1] + eigen$values[2] + eigen$values[3] + eigen$values[4])
}
x <- round(x, digits = 3)
x 
```

## VARIMAX LOADINGS

```{r}
varimax <- varimax(a)
varimax

varimax <- c(varimax$loadings)
varimax <- data.frame(varimax[1:4], varimax[5:8])
```

## FINAL RESULTS IN DATAFRAME 

```{r}
final <- data.frame(a,varimax, com, var)
final <- rbind(final, total.var, x)
final[5,5] <- sum(com)
final[6,5] <- sum(x)
final[5:6,6] <- NA
final[5:6,6] <- NA
final[5,3] <- 2.019
final[5,4] <- 1.932
final[6,3] <- 0.505
final[6,4] <- 0.483


names(final)[names(final) == "X1"] <- "F1"
names(final)[names(final) == "X2"] <- "F2"
names(final)[names(final) == "com"] <- "Communalities"
names(final)[names(final) == "var"] <- "Specific Variances"
names(final)[names(final) == "varimax.1.4."] <- "Varimax F1"
names(final)[names(final) == "varimax.5.8."] <- "Varimax F2"

final <- cbind(Row.Names = c("8 years", "8.5 years", "9 years", "9.5 years", "Variance", "Proportion"), final)
names(final)[names(final) == "Row.Names"] <- "Variables"

final["F1"] <- round(final["F1"],3)
final["F2"] <- round(final["F2"],3)
final["Varimax F1"] <- round(final["Varimax F1"],3)
final["Varimax F2"] <- round(final["Varimax F2"],3)
final
```